The UK Biobank Study: Systematic Review of the Reporting Quality in UK Biobank

### 2021-11-29 manuscript/manuscript.Rmd
fix flow diagram, move code for supplementary material up so all code is in same chuck at start so all objects for in-text r code will exist, small corrections, move deviations from protocol supplementary material, drop supp table 2 & change supp table 1 to figure & update references accordingly, add limitations

---
title: 'The UK Biobank Study: Systematic Review of the Reporting Quality in UK Biobank
  Studies'
author: Emma Anderson, George Davey Smith, Katie Drax, Mark Gibson, Marcus Munafò,
  Benjamin Woolf, and Rebecca Richmond
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    number_sections: yes
  html_document:
    df_print: paged
  pdf_document: default
csl: apa.csl
bibliography: manuscript.bib
---

@@

# Abstract

Background

The STrengthening the Reporting of OBservational studies in Epidemiology (STROBE)
statement provides a reporting guideline for observational studies, including secondary analyses.
We used the STROBE guidelines to assess the reporting quality of studies using UK Biobank data.

Participants

We assessed 70 cohort, cross-sectional and case-control studies published before 2019 using the STROBE guidelines.

Results
We identified ??? studies, included ???, and assessed a random sample of ???.

Conclusions

This study was preregistered before we began to screen the search results [@drax2019].

## Eligibility criteria

Eligible articles were full research articles that reported analyses of UKB data. 
We give the reason for exclusion for each article we excluded at the full-text level of screening.

Inclusion criteria:

* Publication date: After 30 March 2012. 
This is the date the UKB was launched and UKB data was first available for researchers to use.
* Publication type: Full research articles (including simulation and modelling articles).
* Language: English. 
Relevant studies written in another language may be relevant but, to ensure accurate and precise evaluation, the article would require a professional translation which is not feasible given the review’s resources.
* Data source: Article’s findings were produced from analysis(es) of data from the UKB. 
This includes studies which reported findings from other studies alongside findings from UKB data.

Exclusion criteria:

* Article findings were produced from analysis(es) of pooled data from multiple studies.
These are studies which have used data from participants from the UKB and other datasets in the same analysis.
* Meta-analyses; narrative and systematic reviews; preprints; post-prints; replies; letters to the editor; corrections; and any other publication types that are not full research articles.
* Retracted articles
* Studies with inaccessible supplementary material. 
We was to ensure our estimates of reporting completeness included the entire version of record.

The full search strategy for each database is included in Appendix A of the protocol [osf ref]. 
To summarise, the terms "UK Biobank", "UKB", "UKBiobank", and "UKB Resource" were included in the "Title", "Abstract", "Keywords", and "All Fields" fields or equivalent of each database.

We conducted four stages of screening for eligibility.
First, using Endnote, KD removed duplicate results, checked the results for completeness, updated the metadata if necessary, and retrieved the full-texts of all articles.
KD updated the metadata by checking each article's Endnote metadata contained the article's DOI, and adding them if they were missing.
KD checked completeness using a list of published full research articles that have used the UKB Resource, provided by the UKB on 27 November 2018. 
We conducted a systematic search despite having this list because we wished to capture all articles from 2018.
Using the full-texts of the results, KD assessed if results were full research articles, and excluded reviews, corrections, conference abstracts, etc.

Second, we imported the results into Rayyan [@ouzzani2016]. 
BW and KD screened them to determine if they met the inclusion criteria.
If insufficient information was available to determine a paper’s eligibility, we contacted the authors and requested the required information, if it was unobtainable we excluded the article.
Time constraints meant we excluded some articles with insufficient information without contacting the authors.

Third, We classified studies as "traditional epidemiology", "Mendelian randomisation" and "Other", according to the type of data they analysed.
After we piloted these classifications on 20 studies, the classification criteria were:

*	Traditional epidemiology – studies investigating associations between exposures and health outcomes using the UK Biobank data alone. 
*	Mendelian randomisation – studies that conduct Mendelian randomisation analyses on UK Biobank data alone
*	Other – articles that do not meet the ‘Traditional epidemiology’ or ‘MR’ classification criteria described above or include analyses of genetic or imaging data.

BW and KD independently classified articles' abstracts.
They examined full texts if classification was unclear from the abstract alone.
They discussed any conflicts and resolved them by mutual consent if possible, if not resolved after discussion then RR assigned the category.
Articles could be allocated into more than one classification group.
We then excluded all but the traditional epidemiology studies.
The Mendialian randomisation studies were not relevant to this paper but we included it as a classification because the data may be used for one of our author's future projects.

Fourth, we catergorised the 178 studies classified as observational epidemiological studies into those with cohort, case-control and cross-sectional study designs using the Study Design Form described below.
During this process KD also examined the notices attached to the online version of the observational epidemiology articles to identify any retracted articles. 
None were identified.
However, we did identify studies that should be excluded, these studies were:

* Studies incorrectly classified as observational epidemiological studies.
* Studies with inaccessible supplementary material. 
* Studies we did consider to be observational epidemiological studies but could not consider them to have cohort, case-control or cross-sectional designs - for example those using prediction models

Different STROBE items are relevant for cohort, case-control and cross-sectional study designs. 
To ensure coders completed the same STROBE items for each article KD and MG independently identified the study designs used by each of the 178 “observational epidemiology” articles by completing the Study Design Form for each article.
In the form, coders also indicated if all supplementary material can be accessed via University of Bristol subscriptions. 
Any conflicts were discussed by the two coders and resolved by mutual consent if possible.
If not resolved after discussion, RR determined the study designs. 
The Study Design Form was hosted on Qualtrics but a PDF version can be found at https://osf.io/jfk24/.

The data extraction form was hosted on Qualtrics. 
This form combined the data items described as part of the "Data Extraction Form" and "Reporting Quality Assessment" in the protocol, but not all were kept.
We conducted two pilots.
In the first, MG, KD and/or RR piloted 5 observational epidemiology articles.
We made major changes to the Data Extraction Form so we completed it again for the articles in the first pilot. 
In the second, KD piloted the changed form on 3 articles. 
We made no major changes so the we retained the responses for articles in the second pilot.
See the PDF version of the data extraction form for full information on our variables and how we measured them at https://osf.io/hngcj/.

The coders independently completed the data extraction form for each article assigned to them.
Almost all variables were categorical so coders assigned the most appropriate category for each variable.
Variables derived from the STROBE items had the same six categories:

Using the data extraction form on Qualtrics, we collected data on the reporting quality on numerous variables, including those derived from the STROBE items.
This mean our raw data estimated the reporting quality for all subdivisions of the STROBE items and for the “starred” items associated with any STROBE items that had an asterisk (*) next to it. 
For the final analysis, we excluded all starred subdivisions because inter-rater reliability was so poor, this was primarily because raters could not agree when the items were or were not applicable.
In the data extraction form raters could indicate items as "NA", "Unsure", "No", "Partially", "Partially-External" or "Yes".
After resolving conflicts, no "Unsure" ratings remained. 
We collapsed the "Partially-External" into the "Yes" catergory.
This was because we agreed reporting there were very few "Partially-External" categories after resolving conflicts, to make the kappa statistics more interpretable, and we understood that reporting where the required information can be found could be considered complete reporting.

Some STROBE items were written as multiple-barrelled questions e.g " Report assessment methods for outcomes, exposures, predictors, confounder, and effect modifiers". 
We subdivided these questions into separate variables in the data extraction form.
We could have assessed each item without dividing it up but collecting the reporting completeness at the subdivision level we have a more granular understanding of what information is or is not reported.
We summed the subdivisions back together to calculate the reporting completeness of each STROBE item they corresponded to. 
This gave us an understanding of the reporting quality for the entire STROBE item and reduced the number of variables from `r length(unique(stab2$strobe_item))` to `r length(unique(bar_data$strobe_item))`.

Figure 1 presents the reporting completeness for each STROBE item, to ensure a comprehensive understanding of our results, Supplementary Table 2 provides the reporting completeness for each subdivision.

```{r table2, echo=F, include = F}

###################
# frequencies non-strobe ####
##################

@@

```
# Discussion

@@

Other items that were poorly reported carry different concerns to those of reproducibility.
Most authors partially reported item 16a. 
Supplementary Table 2 shows that while nearly all authors fully report their confounder-adjusted estimates most are failing report their unadjusted elements and to justify why they included the confounders they did.  
This information may be less vital to reproducing analyses than other items, but it does suggest a widespread norm of researchers including confounders without a stated justification.
Similarly the significant lack of reporting for item 9, 19 and 22 suggests it is acceptable for researchers not to provide detailed analysis of potential bias in their studies or report the role of funders in their studies.
Lastly, many authors failing to mention the design of their study in their title or abstract (item 1a) could create problems in retrieving studies during searches.

@@

* appropriateness of strobe for secondary data analysis - e.g. item 10 excluded completely
* kappa statistics -- limitations of double coding for something so subjective
* deviations from systematic review method & protocol
* reverse engineering of reporting guidelines = inappropriate to use as reporting assessment?
  ** Releasing ‘adherence checklists’ with reporting guidelines would help 
reporting assessments

@@

@@

```{r STable1, echo= F}

create_stab1 <- function(){
  # read in rules to add into table
  stab1 <- read.csv("../data/rules.csv", encoding = "UTF-8", na.strings = c(""), stringsAsFactors = F)

  # save ids of all variable cols
  cols_id <- grep("variable", colnames(stab1))

  # save variables element so can save for loop products
  x <- c()

  # paste all variable cols together and save products
  for(i in 1:nrow(stab1)){
    x <- c(x, paste(stab1[i, cols_id], collapse = ","))
  }

  # clean variable names to strobe items and remove all missing items and spaces
  x <- gsub(",NA.*| |starred|^s", "", x) %>%
    gsub(",s", ",", .) %>%
    # add spaces back in so more readable
    gsub(",", ", ", .)

  # add in as column
  stab1$STROBEItem <- x

  colnames(stab1)[colnames(stab1) == "rule"] <- "Rule"
  colnames(stab1)[colnames(stab1) == "STROBEItem"] <- "STROBE Item" 
  # remove all unnecessary cols
  stab1 <- stab1[, -grep("variable|response", colnames(stab1))]

  stab1$Rule <- gsub("^Rule = |^Rule =", "", stab1$Rule)

  return(stab1)

}

stab1 <- create_stab1()

pander(stab1, caption = "Supplementary Table 1. Rules for resolving conflicts", justify = c("left", "center"))
```

## Supplementary Table 2

```{r STable2, echo=F}

  # remove s from strobe item
  stab2$strobe_item <- gsub("^s|starred", "", stab2$strobe_item)

# strobe item on x axis and plot % yes
stab2 %>%ggplot(aes(x=strobe_item, y=..count.., fill = value)) + 
  # blue fill
  geom_bar() +
  theme_classic() +
  theme(text = element_text(size=30))+
  #scale_x_discrete(labels= labels$x_labels) +
  #scale_y_continuous(breaks = seq(0, nrow(df), by = 10)) + 
  coord_flip() +
  xlab("STROBE Item") +
  ylab("Count") +
  scale_fill_brewer(palette = "Greys") +
  #geom_text(aes(label = count), vjust = 0, size = 8) +
  theme(plot.title = element_text(size = 22)) 
dev.off()

pander(stab2_tab, caption = "Supplementary Table 2. Descriptive statistics for reporting quality of subdivisions of all STROBE items")
```

## Supplementary Table 3

```{r, echo = F}
pander(kappa_sub, caption = "Supplementary Table 3. Kappa statistics for reporting quality of subdivisions of all STROBE items")
```

>>>>

---
title: 'The UK Biobank Study: Systematic Review of the Reporting Quality in UK Biobank
  Studies'
author: Emma Anderson, George Davey Smith, Katie Drax, Mark Gibson, Marcus Munafò, Robert Thibault, Benjamin Woolf, and Rebecca Richmond
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    number_sections: yes
  html_document:
    df_print: paged
  pdf_document: default
csl: apa.csl
bibliography: manuscript.bib
---

@@

kappa_item$STROBEItem <-   gsub("_sum|starred|^s", "", kappa_item$STROBEItem)

###################
# create table 2 - frequencies non-strobe ####
##################

@@

#############
# create stab2####
############

labels <-  read.csv("../outputs/extraction_dict.csv", encoding = "UTF-8", na.strings = c(""), stringsAsFactors = T)

stab2 <- left_join(stab2, labels, by = c("strobe_item" = "variable"))

stab2$question <- gsub("Indicate if the authors report the following items for all the studies in the article \\(unless a study design is specified\\) - |Indicate if the authors report the following items for all the cross-sectional studies in the article - |_ - |Indicate if the authors report the following items for all the case-control studies in the article - |Indicate if the authors report the following items for all the cohort studies in the article - |Indicate if the authors report the following items for all the cross-sectional and cohort studies in the article - ", "", stab2$question)

# strobe item on x axis and plot % yes
stab2 %>%ggplot(aes(x=strobe_item, y=..count.., fill = value)) + 
  # blue fill
  geom_bar() +
  theme_classic() +
  theme(text = element_text(size=30))+
 # scale_x_discrete(labels= unique(stab2$question)) +
  #scale_y_continuous(breaks = seq(0, nrow(df), by = 10)) + 
  coord_flip() +
  xlab("STROBE Item") +
  ylab("Count") +
  scale_fill_brewer(palette = "Greys") +
  #geom_text(aes(label = count), vjust = 0, size = 8) +
  theme(plot.title = element_text(size = 22)) 
dev.off()

``` {r}
# Abstract
#
#Background
#
#The STrengthening the Reporting of OBservational studies in #Epidemiology (STROBE)
#statement provides a reporting guideline for observational #studies, including secondary analyses.
#We used the STROBE guidelines to assess the reporting quality #of studies using UK Biobank data.
#
#Participants
#
#We assessed 70 cohort, cross-sectional and case-control #studies published before 2019 using the STROBE guidelines.
#
#Results
#We identified ??? studies, included ???, and assessed a #random sample of ???.
#
#Conclusions
```

This study was preregistered before we began to screen the search results [@drax2019].
For a description of the deviations from the registered protocol see the Supplementary Material.

## Inclusion criteria

The inclusion criteria were:

* Publication date: After 30 March 2012. 
This is the date the UKB was launched and UKB data was first available for researchers to use.
* Publication type: Full research articles (including simulation and modelling articles).
* Language: English. 
Relevant studies written in another language may be relevant but, to ensure accurate and precise evaluation, the article would require a professional translation which is not feasible given the review’s resources.
* Data source: Article’s findings were produced from analysis(es) of data from the UKB. 
This includes studies which reported findings from other studies alongside findings from UKB data.
* Study type: Observational epidemiology studies
* Study design: Cross-sectional, cohort, or case-control study design.

The exclusion criteria were:

* Article findings were produced from analysis(es) of pooled data from multiple studies.
These are studies which have used data from participants from the UKB and other datasets in the same analysis.
* Meta-analyses; narrative and systematic reviews; preprints; post-prints; replies; letters to the editor; corrections; and any other publication types that are not full research articles.
* Retracted articles
* Studies with inaccessible supplementary material. 
This was to ensure our estimates of reporting completeness included the entire version of record.

The full search strategy for each database is included in Appendix A of the protocol [@drax2019]. 
To summarise, the terms "UK Biobank", "UKB", "UKBiobank", and "UKB Resource" were included in the "Title", "Abstract", "Keywords", and "All Fields" fields or equivalent of each database.

We conducted three stages of screening for eligibility.
First, using Endnote, KD removed duplicate results, checked the results for completeness, updated the metadata if necessary, and retrieved the full-texts of all articles.
KD updated the metadata by checking each article's Endnote metadata contained the article's DOI, and adding them if they were missing.
KD checked completeness using a list of published full research articles that have used the UKB Resource, provided by the UKB on 27 November 2018. 
We conducted a systematic search despite having this list because we wished to capture all articles from 2018.
Using the full-texts of the results, KD assessed if results were full research articles, and excluded reviews, corrections, conference abstracts, etc.

Second, we imported the results into Rayyan [@ouzzani2016]. 
BW and KD independently screened the results and indicated if they met the inclusion criteria.
At this point we included all articles whether or not they had a cross-sectional, cohort, or case-control study design because the "Study design" inclusion critera was added later.
They examined full texts if inclusion was unclear from the abstract alone.
They discussed any conflicts and resolved them by mutual consent if possible, if not resolved after discussion then RR assigned the category.
If insufficient information was available in the full text to determine a paper’s eligibility, we contacted the authors and requested the required information, if it was unobtainable we excluded the article.
Time constraints meant we excluded XXX articles with insufficient information without contacting the authors.

Different STROBE items are relevant for cohort, case-control and cross-sectional study designs. 
To ensure coders completed the same STROBE items for each article KD and MG independently identified the study designs used by each of the 178 “observational epidemiology” articles by completing the Study Design Form for each article.
In the form, coders also indicated if all supplementary material can be accessed via University of Bristol subscriptions. 
Any conflicts were discussed by the two coders and resolved by mutual consent if possible.
If not resolved after discussion, RR determined the study designs. 
The Study Design Form was hosted on Qualtrics but a PDF version can be found at https://osf.io/jfk24/.
During this process we also identified studies we did consider to be observational epidemiological studies but could not consider them to have cohort, case-control or cross-sectional designs - for example those using prediction models.
Any articles we could not classify as cohort, case-control, or cross-sectional were excluded at this point.

Two coders independently completed the data extraction form for each article assigned to them.
Almost all variables in the form were categorical so coders assigned the most appropriate category for each variable.
We conducted two pilots.
In the first, MG, KD and/or RR piloted 5 observational epidemiology articles.
We made major changes to the Data Extraction Form so we completed it again for the articles in the first pilot. 
In the second, KD piloted the changed form on 3 articles. 
We made no major changes so we kept the responses for articles in the second pilot.
The PDF version of the data extraction form contains full information on our variables and how we measured them and is available at https://osf.io/hngcj/.

The form contained questions to measure variables relating to the reporting quality of different information.
We derived some variables from the STROBE guidelines.
Some STROBE items were written as multiple-barrelled questions e.g " Report assessment methods for outcomes, exposures, predictors, confounder, and effect modifiers". 
We subdivided these questions into separate variables in the data extraction form.
We could have assessed each item without dividing it up but collecting the reporting completeness at the subdivision level we have a more granular understanding of what information is or is not reported.

Variables derived from the STROBE items had the same six categories:

Using the data extraction form on Qualtrics, we collected data on the reporting quality on numerous variables, including those derived from the STROBE items.
This mean our raw data estimated the reporting quality for all subdivisions of the STROBE items and for the “starred” items associated with any STROBE items that had an asterisk (*) next to it. 
For the final analysis, we excluded all starred subdivisions because inter-rater reliability was so poor, this was primarily because raters could not agree when the items were or were not applicable.
In the data extraction form raters could indicate items as "NA", "Unsure", "No", "Partially", "Partially-External" or "Yes".
After resolving conflicts, no "Unsure" ratings remained. 
We collapsed the "Partially-External" into the "Yes" catergory.
This was because we agreed reporting there were very few "Partially-External" categories after resolving conflicts, to make the kappa statistics more interpretable, and we understood that reporting where the required information can be found could be considered complete reporting.
We summed the subdivisions back together to calculate the reporting completeness of each STROBE item they corresponded to. 
This gave us an understanding of the reporting quality for the entire STROBE item and reduced the number of variables from `r length(unique(stab2$strobe_item))` to `r length(unique(bar_data$strobe_item))`.

Figure 1 presents the reporting completeness for each STROBE item, to ensure a comprehensive understanding of our results, Supplementary Figure 1 provides the reporting completeness for each subdivision.

# Discussion

@@

Other items that were poorly reported carry different concerns to those of reproducibility.
Most authors partially reported item 16a. 
Supplementary Figure 1 shows that while nearly all authors fully report their confounder-adjusted estimates most are failing report their unadjusted elements and to justify why they included the confounders they did.  
This information may be less vital to reproducing analyses than other items, but it does suggest a widespread norm of researchers including confounders without a stated justification.
Similarly the significant lack of reporting for item 9, 19 and 22 suggests it is acceptable for researchers not to provide detailed analysis of potential bias in their studies or report the role of funders in their studies.
Lastly, many authors failing to mention the design of their study in their title or abstract (item 1a) could create problems in retrieving studies during searches.

@@

Our study has several limitations.
Firstly, many variables in the data extraction form we used to assess reporting quality had poor inter-rater reliability.
The kappa statistics, use of "Rules" to solve conflicts, and exclusion of the starred subdivisions suggest as much.
This questions the appropriateness of our data extraction form, and the STROBE guidelines it is a based on, as a measure of reporting quality since it does not seem to accurately estimate reporting quality across coders.
This poor reliability may be because the STROBE guidelines were not designed as a measure of reporting quality.
The STROBE guidelines are reporting guidelines designed for authors to use to check their manuscript contains all the relevant information, while they are writing up.
The guidelines are not designed as an assessment of reporting by people other than the authors.
Reverse engineering the guidelines for a purpose they were not designed for may be why the inter-rater reliability is poor.
The appropriateness of the STROBE guidelines for secondary data analyses, such as those using the UK Biobank, is also questionable.
We used the STROBE guidelines to create our data extraction form because it was the most appropriate reporting guidelines available.
However, the STROBE guidelines did not necessarily capture the information we needed, leading to us excluding several STROBE items and including additional variables.
For example, the guidelines were written in 2007, therefore there is no question on whether or how authors reported analysis code.
We also excluded several items or subdivisions of items because they requested reporting of the same information or information not applicable to a seondary data analysis, specifically item 10, two subdivisions of item 6a, one subdivision of item 5, and one subdivision of item 8.
See the rule guidelines for further information on these duplicate and inapplicable variables (https://osf.io/hngcj/).

@@

```{r}
#Third, using Rayyan again, we classified studies as "traditional epidemiology", "Mendelian randomisation" and "Other", according to the #type of data they analysed.
#After we piloted these classifications on 20 studies, the classification criteria were:
#
#*	Traditional epidemiology – studies investigating associations between exposures and health outcomes using the UK Biobank data alone. 
#*	Mendelian randomisation – studies that conduct Mendelian randomisation analyses on UK Biobank data alone
#*	Other – articles that do not meet the ‘Traditional epidemiology’ or ‘MR’ classification criteria described above or include analyses of #genetic or imaging data.
#
#BW and KD independently classified articles' abstracts.
#They examined full texts if classification was unclear from the abstract alone.
#They discussed any conflicts and resolved them by mutual consent if possible, if not resolved after discussion then RR assigned the #category.
#Articles could be allocated into more than one classification group.
#We then excluded all but the traditional epidemiology studies.
#The Mendialian randomisation studies were not relevant to this paper but we included it as a classification because the data may be used for one of our author's future projects.

The final data extraction form combined the data items described as part of the "Data Extraction Form" and "Reporting Quality Assessment" in the protocol, but not all were kept.

## Supplementary Figure 1

*Supplementary Figure 1. Descriptive statistics for reporting quality of subdivisions of all STROBE items*
![](Rplot10.png)

@@

```{r, echo = F}
pander(kappa_sub, caption = "Supplementary Table 1. Kappa statistics for reporting quality of subdivisions of all STROBE items")
```
